---
title: "Introduction to Graphs"
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
---


```{r setup, include=FALSE}
      library(ggraph)
      library(tidygraph)
      library(igraph)
```

## What are graphs?

- A graph $G$ is a set of two sets

  - A set of nodes $V$ (points, vertices)
  - A set of edges $E$ (lines, links)
  
$$
  G = \{E, V\} 
$${#eq-graph1}

- Nodes represent the entities (e.g., actors) in the social network
- Edges represent the relations or connections between entities

## What are graphs?

:::: {.columns}
::: {.column width="45%"}
- On the right you can see (a pictorial representation of) a graph
  - The graph's node set is:
    - $V = \{A, B, C, D\}$
  - The graph's edge set is:
    - $E = \{AB, AC\}$
- The cardinality of the node set is four $|V| = 4$
- The cardinality of the edge set is two $|E| = 2$
:::

::: {.column width="55%"}
```{r}
    gr <- create_empty(4) %>% 
        bind_edges(data.frame(from = 1, to = 2:3)) %>% 
        mutate(name = toupper(letters[1:4])) 
    p <- ggraph(gr, layout = 'grid') 
    p <- p + geom_edge_link(color = "steelblue", edge_width = 1.25) 
    p <- p + geom_node_point(aes(x = x, y = y), size = 22, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 12, color = "white")
    p <- p + theme_graph()
    p <- p + scale_x_continuous(expand = expand_scale(c(.10, .10))) 
    p <- p + scale_y_continuous(expand = expand_scale(c(.10, .10)))
    p
```
:::
::::

## What are graphs?

:::: {.columns}
::: {.column width="45%"}
- We can refer to graphs by their number of nodes and edges.
  - Like this: $G(|V|, |E|)$
- The graph on the right is a $G(4, 2)$ graph
  - It has four nodes and two edges
- The $G(1, 0)$ graph is called the **trivial graph**
  - One is the loneliest number!
- The smallest non-trivial graph is the $G(2, 1)$ graph
  - Also referred to as a **connected dyad**

:::

::: {.column width="55%"}
```{r}
    p
```

:::
::::
```{r}

```
