---
title: "Homework V: Affiliation Networks"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
    require(kableExtra)
```
```{r Southern women data}
    A <- matrix(
         c(
          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0, 
          
          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0, 
          
          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
          ncol = 14, byrow = TRUE)
     p <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
     p.abv <- c("EV", "LA", "TH", "BR", "CH", "FR", "EL", "RU", "VE", "MY", "KA", "SY", "NO", "HE", "OL", "FL", "PE", "DO")
     g <- paste("E", c(1:14), sep = "")
     rownames(A) <- p
     colnames(A) <- g
     set.seed(682)
     A <- A[sample(1:18, 10), sample(1:14, 8)]
     affil.el <- data.frame(matrix(ncol = 2, nrow = 0))
          k <- 1
          for (i in 1:nrow(A)) {
              for(j in 1:ncol(A)) {
                  if (A[i, j] == 1) {
                    affil.el[k, 1] = rownames(A)[i]
                    affil.el[k, 2] = colnames(A)[j]
                  k <- k + 1
                  }
              }
          }
```

```{r}
#| label: fig-grex1
#| fig-cap: "Southern women bipartite graph."
#| fig-cap-location: margin   
#| fig-height: 8
#| fig-width: 12

    gr <- graph_from_data_frame(data.frame(affil.el), directed = FALSE)
    V(gr)$type <- bipartite_mapping(gr)$type
    p <- ggraph(gr, layout = 'bipartite') 
    p <- p + geom_edge_link(color = "steelblue", edge_width = 1.1) 
    p <- p + geom_node_point(color = c(rep("white", 10), rep("tan2", 8)), 
                             shape = c(rep("square", 10), rep("circle", 8)), 
                             size = c(rep(15, 10), rep(20, 8)))
    p <- p + geom_node_text(aes(label = c(rownames(A), colnames(A))), 
                            size = c(rep(5, 10), rep(6, 8)), 
                            color = "black")   
    p <- p + theme_graph() + theme(legend.position = "none") 
    p
```

The Figure above shows the **bipartite graph** corresponding to (part of) the *Southern Women Events Data* a classic social network data set collected in the 1930s by the sociologist Allison Davis and collaborators and included in book [*Deep South*](https://www.amazon.com/Deep-South-Anthropological-Southern-Classics/dp/1570038155). The graph represents observed attendance at 14 social events by 18 Southern "high society" women from the South. 

1. Write down the **affiliation matrix** corresponding to the Southern women network bipartite graph.

```{r}
    a <- matrix(" ", nrow = 10, ncol = 8)
    rownames(a) <- rownames(A)
    colnames(a) <- colnames(A)
    kbl(a, format = "html", padding = 1, align = "c") %>% 
        kable_classic(lightable_options = "striped") %>% 
        row_spec(1:10, color = "black", extra_css = "border-bottom: 1px solid;") %>% 
        column_spec(1:8, color = "black", border_right = TRUE, border_left = TRUE)
```

1. Which women attend the most events?
<br>
<br>
<br>

1. Which women attend the least events?
<br>
<br>
<br>

1. What is the average number of events attended by the women?
<br>
<br>
<br>

1. Which are the largest event(s)?
<br>
<br>
<br>

1. Which are the smallest event(s)?
<br>
<br>
<br>

1. What is the average event size?
<br>
<br>
<br>

1. Write down the *transpose* of the **affiliation matrix** corresponding to the Southern women network bipartite graph.

```{r}
    t.a <- matrix(" ", nrow = 8, ncol = 10)
    rownames(t.a) <- colnames(A)
    colnames(t.a) <- paste(substr(rownames(A) , start = 1 , stop = 3), ".", sep = "")
    kbl(t.a, format = "html", padding = 1, align = "c") %>% 
        kable_classic(lightable_options = "striped") %>% 
        row_spec(1:8, color = "black", extra_css = "border-bottom: 1px solid;") %>% 
        column_spec(1:10, color = "black", border_right = TRUE, border_left = TRUE)
```

Using the information in 1 and 2 and what you know about **matrix multiplication** (5pts):

1. Write down the entries corresponding to the **co-membership matrix** of the Southern women.

```{r}
    c.m <- matrix(" ", nrow = 10, ncol = 10)
    rownames(c.m) <- paste(substr(rownames(A) , start = 1 , stop = 3), ".", sep = "")
    colnames(c.m) <- paste(substr(rownames(A) , start = 1 , stop = 3), ".", sep = "")
    kbl(c.m, format = "html", padding = 1, align = "c") %>% 
        kable_classic(lightable_options = "striped") %>% 
        row_spec(1:10, color = "black", extra_css = "border-bottom: 1px solid;") %>% 
        column_spec(1:10, color = "black", border_right = TRUE, border_left = TRUE)
```

1. Write down the entries corresponding to the **group overlap matrix** of the Southern women.

```{r}
    g.o <- matrix(" ", nrow = 8, ncol = 8)
    rownames(g.o) <- colnames(A)
    colnames(g.o) <- colnames(A)
    kbl(g.o, format = "html", padding = 1, align = "c") %>% 
        kable_classic(lightable_options = "striped") %>% 
        row_spec(1:8, color = "black", extra_css = "border-bottom: 1px solid;") %>% 
        column_spec(1:8, color = "black", border_right = TRUE, border_left = TRUE)
```