---
title: "Homework III: Graph Metrics"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
    require(kableExtra)
```

```{r}
    d <- c(2, 5, 4, 5, 2, 3, 6, 2, 2, 6, 4, 1, 3, 3, 1, 3)
```

Consider the graph shown in Figure 1:

1. What is the **order** of the graph?
<br>
`r length(d)`

1. Write down the graph's **degree sequence**:
<br>
`r as.matrix(sort(d, decreasing = TRUE))`

1. What is the graph's **degree range**?
<br>
`r max(d) - min(d)`

1. What is the graph's **sum of degrees**?
<br>
`r sum(d)`

1. What is the **size** of the graph?
<br>
`r sum(d)/2`

1. What is the graph's **average degree**?
<br>
`r sum(d)/length(d)`

1. What is the graph's **maximum size**?
<br>
`r (length(d) * (length(d) - 1))/2`

1. Compute the **density** of the graph:
<br>
`r round((sum(d)/2)/((length(d) * (length(d) - 1))/2), 2) `
<br>
`r round((sum(d)/2)/((length(d) * (length(d) - 1))/2), 3) `
<br>
`r round((sum(d)/2)/((length(d) * (length(d) - 1))/2), 4) `

```{r}
#| label: fig-grex2
#| fig-cap: "A directed graph."
#| fig-cap-location: margin   
#| fig-height: 12
#| fig-width: 12

    set.seed(924)
    gr <- play_gnm(n = 13, m = 28, directed = TRUE, loops = FALSE)
    gr <- mutate(gr, name = toupper(letters[1:13]))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_parallel(color = "steelblue", edge_width = 1.1,
                            arrow = arrow(length = unit(4, 'mm')),
                            end_cap = circle(7, 'mm'), 
                            sep = unit(8, 'mm'))    
    p <- p + geom_node_point(aes(x = x, y = y), size = 16, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 7, color = "white")
    p <- p + theme_graph()
    out.d <- degree(gr, mode = "out")
    in.d <- degree(gr, mode = "in")
```

Consider the graph shown in Figure 2:

1. What is the graph's **out-degree sequence**?
<br>
`r sort(as.vector(out.d), decreasing = TRUE)`

1. What is the graph's **in-degree sequence**?
<br>
`r sort(as.vector(in.d), decreasing = TRUE)`

1. What is the graph's **out-degree range**?
<br>
`r max(out.d) - min(out.d)`

1. What is the graph's **in-degree range**?
<br>
`r max(in.d) - min(in.d)`

1. What is the graph's **sum of degrees**?
<br>
`r sum(in.d)`

1. What is the **size** of the graph?
<br>
`r sum(in.d)`

1. What is the graph's **maximum size**?
<br>
`r length(in.d) * (length(in.d) - 1)`

1. Compute the **density** of the graph:
<br>
`r round(sum(in.d) / (length(in.d) * (length(in.d) - 1)), 2)`
<br>
`r round(sum(in.d) / (length(in.d) * (length(in.d) - 1)), 3)`
<br>
`r round(sum(in.d) / (length(in.d) * (length(in.d) - 1)), 4)`
