---
title: "Homework I: Graph Theory"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
```

## Vertex and edge sets

```{r}
#| label: fig-grex1
#| fig-cap: "A simple graph."
#| fig-cap-location: margin   
#| fig-height: 8
#| fig-width: 8
    set.seed(529)
    n <- 8
    gr <- play_gnm(n = n, m = 16, directed = FALSE, loops = FALSE)
    gr <- mutate(gr, name = toupper(letters[1:n]))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.25) 
    p <- p + geom_node_point(aes(x = x, y = y), size = 18, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 8, color = "white")
    p <- p + theme_graph() + ylim(-1, 1)
    p
```

Consider the graph shown in @fig-grex1:

1. Write down the **vertex set** of the graph: 
<br>
<br>
<br>

1. Write down the **edge set** of the graph:
<br>
<br>
<br>

## Node Neighborhoods


```{r}
#| label: fig-grex2
#| fig-cap: "Another simple graph."
#| fig-cap-location: margin 
#| fig-height: 12
#| fig-width: 12
    set.seed(471)
    n <- 12
    gr <- play_gnm(n = n, m = 20, directed = FALSE, loops = FALSE)
    gr <- mutate(gr, name = toupper(letters[1:n]))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.25) 
    p <- p + geom_node_point(aes(x = x, y = y), size = 18, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 8, color = "white")
    p <- p + theme_graph() 
    p
    
```

Consider the graph shown in @fig-grex2:

1. Write down the **neighborhood** of node *D*
<br>
<br>
<br>

1. Write down the **neighborhood** of node *K*
<br>
<br>
<br>

1. What is the **intersection** of the neighborhoods of nodes *D* and *I*?
<br>
<br>
<br>

1. What is the **intersection** of the neighborhoods of nodes *E* and *F*?
<br>
<br>
<br>

1. What is the **union** of the neighborhoods of nodes *H* and *J*? 
<br>
<br>
<br>

## Node Degree

Consider the graph shown in @fig-grex2:

1. What is the **degree** of node *B*?
<br>
<br>

1. What is(are) the node(s) with the largest **degree**?
<br>
<br>

1. What is(are) the node(s) with the smallest **degree**?
<br>
<br>

## Subgraphs

1. Go back to @fig-grex1. Draw the **node-deleted** subgraph of this graph that *excludes* nodes *D* and *G*
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
