---
title: "Homework VII: Subgroups"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
    require(kableExtra)
```

```{r}
#| label: clique
#| fig-width: 12
#| fig-height: 12
#| fig-cap: "An undirected graph."
#| fig-cap-location: margin

    set.seed(693)
    gr1 <- create_complete(4) %>% 
        mutate(name = toupper(letters[1:4]))
    gr2 <- create_complete(4) %>% 
        mutate(name = toupper(letters[5:8]))
    gr3 <- create_complete(5) %>% 
        mutate(name = toupper(letters[9:13]))
    gr4 <- play_islands(n_islands = 4, size_islands = 4, 
                        p_within = 1, m_between = 1) %>% 
         mutate(name = toupper(letters[1:16]))
    gr <- bind_graphs(gr1) %>% 
         bind_graphs(gr2) %>% 
         bind_graphs(gr3) %>% 
         graph_join(gr4)
    p <- ggraph(gr, layout = 'kk') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.15) 
    p <- p + geom_node_point(color = "tan2", size = 16) 
    p <- p + geom_node_text(aes(label = name), size = 8, color = "white")
    p <- p + theme_graph() 
    p
```
1. List the nodes in all the **cliques** of *size four* in the graph.
<br>
<br>
```{r}
    c <- cliques(gr, min = 4, max = 4)
    kbl(toupper(matrix(letters[unlist(c)], ncol = 4, byrow = TRUE)))

```

  
1. How many **cliques** of *size four* does node *M* belong to?
<br>
5

1. List the edges would have to be added to the graph to make nodes $\{A, C, F, J, O\}$ into a clique of size 5:
<br>
```{r}
    s <- induced_subgraph(gr, c("A", "C", "F", "J", "O"))
    cs <- complementer(s)
    e <- as_edgelist(cs)
    kbl(e)
```

1. Write down a set of *two* nodes that if removed would disconnect the graph:
<br>
EF

1. Write down a set of *three* edges that if removed would disconnect the graph:
<br>
CF, AO, BP
<br>
FJ, EO, FC

1. List the nodes in two 2-cliques in the graph:
<br>
[Generously grade]

1. List the nodes in one 3-clique in the graph:
<br>
[Generously grade]


1. Imagine we removed the edges $\{FJ, FC, EO, AO, BP\}$. How many components would the resulting graph have?
<br>
```{r}
#| fig-width: 12
#| fig-height: 12

    gr2 <- gr - edge("F|J")
    gr2 <- gr2 - edge("C|F")
    gr2 <- gr2 - edge("E|O")
    gr2 <- gr2 - edge("B|P")
    gr2 <- gr2 - edge("A|I")

    p <- ggraph(gr2, layout = 'kk') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.15) 
    p <- p + geom_node_point(color = "tan2", size = 16) 
    p <- p + geom_node_text(aes(label = name), size = 8, color = "white")
    p <- p + theme_graph() 
    p
```


1. List the nodes in the *five cliques of size four* nested inside the clique of size five.
<br>
<br>

```{r}
    c <- cliques(gr, min = 4, max = 4)
    m <- matrix(LETTERS[unlist(c)], ncol = 4, byrow = TRUE)
    kbl(m[-c(2, 3, 4), ])
```

1. List the nodes in the *four cliques of size three* nested inside one of the cliques of size four.
<br>
Example:

```{r}
    s <- induced_subgraph(gr, c("G", "F", "H", "E"))
    c <- cliques(s, min = 3, max = 3)
    kbl(toupper(matrix(letters[unlist(c)], ncol = 3, byrow = TRUE)))
```