---
title: "Homework II: Directed Graphs"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
    require(kableExtra)
```

```{r}
#| label: fig-grex1
#| fig-cap: "A directed graph."
#| fig-cap-location: margin   
#| fig-height: 12
#| fig-width: 12
    set.seed(528)
    gr <- play_gnm(n = 10, m = 26, directed = TRUE, loops = FALSE)
    gr <- mutate(gr, name = toupper(letters[1:10]))
    gr <- gr - edge("B|D")
    gr <- gr - edge("I|H")
    gr <- gr - edge("F|H")
    gr <- gr - edge("F|C")
    gr <- gr - edge("G|C")
    gr <- gr - edge("A|G")
    gr <- gr - edge("J|A")
    gr <- add_edges(gr, c(9, 10, 4,5, 4,8, 9,4, 1,4, 9,5))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_parallel(color = "steelblue", edge_width = 1.25,
                            arrow = arrow(length = unit(5, 'mm')),
                            end_cap = circle(10, 'mm'), 
                            sep = unit(8, 'mm'))    
    p <- p + geom_node_point(aes(x = x, y = y), size = 22, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 10, color = "white")
    p <- p + theme_graph()
    p
```



Consider the graph shown in @fig-grex1:

- Write down the **edge set** of the graph:
```{r}
    b <- as_edgelist(gr)
    kbl(b)
```


- Who are the **out-neighbors** of node *A*?
```{r}
   a.out <- t(rownames(as.matrix(neighbors(gr, "A", mode = "out"))))
    kbl(a.out)
```


- Who are the **in-neighbors** of node *D*?
```{r}
   d.in <- t(rownames(as.matrix(neighbors(gr, "D", mode = "in"))))
    kbl(d.in)
```


- What is the **intersection** of the **out-neighborhoods** of nodes *D* and *A*?
```{r}
   d.out <- t(rownames(as.matrix(neighbors(gr, "D", mode = "out"))))
    kbl(t(intersect(a.out, d.out)))
```

- What is the **intersection** of the **in-neighborhoods** of nodes *D* and *F*?
```{r}
   f.in <- t(rownames(as.matrix(neighbors(gr, "F", mode = "in"))))
    kbl(t(intersect(f.in, d.in)))
```

- What is the **union** of the **out-neighborhoods** of nodes *E* and *A*?
```{r}
   a.out <- t(rownames(as.matrix(neighbors(gr, "A", mode = "out"))))
   e.out <- t(rownames(as.matrix(neighbors(gr, "E", mode = "out"))))
    kbl(t(union(a.out, e.out)))
```

- What is the **union** of the **in-neighborhoods** of nodes *D* and *F*?
```{r}
    kbl(t(union(d.in, f.in)))
```

- What is the **out-degree** of node *E*?
<br>
$k(E)_{out}$ = `r length(neighbors(gr, "E", mode = "out"))`

- What is the **in-degree** of node *D*?
<br>
$k(D)_{in}$ = `r length(neighbors(gr, "D", mode = "in"))`


- What is(are) the node(s) with the largest **out-degree**?
```{r}
    d <- as.matrix(degree(gr, mode = "out"))
    kbl(t(d), format = "html") %>% 
    kable_styling()
```
<br>
A D

1. What is(are) the node(s) with the largest **in-degree**?
```{r}
    d <- as.matrix(degree(gr, mode = "in"))
    kbl(t(d), format = "html") %>% 
        kable_styling()
```
<br>
F I

1. What is(are) the node(s) with the smallest **out-degree**?
<br>
B F G

1. What is(are) the node(s) with the smallest **in-degree**?
<br>
A C


1. Which is(are) the **receiver node(s)** in the graph?
<br>
B F G

1. Which is(are) the **transmitter node(s)** in the graph?
<br>
A C

1. Write down the set of **mutual dyads** in the graph:
<br>
ED or DE <br>
DH or HD <br>
IJ or JI <br>
DI or ID <br>
EI or IE <br>

Mutuals Matrix: <br>

```{r}
    a <- as.matrix(as_adjacency_matrix(gr))
    b <- a * t(a)
    kbl(b, format = "html") %>% 
        kable_classic()
```

1. What type of **dyad** is formed by the subgraph containing nodes *C* and *F*?
<br>
Asymmetric or Non-reciprocal

1. What type of **dyad** is formed by the subgraph containing nodes *B* and *C*?
<br>
Null or disconnected

1. What type of **dyad** is formed by the subgraph containing nodes *E* and *I*?
<br>
Mutual or Reciprocal

1. How many **mutual dyads** are there in the graph?
<br>
Five (5)




