---
title: "Homework VIII: Graph Connectivity and Structural Similarity"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    library(ggraph)
    library(tidygraph)
    library(igraph)
    library(kableExtra)
```


```{r}
#| label: fig-grex1
#| fig-cap: "An undirected graph."
#| fig-cap-location: margin   
#| fig-height: 8
#| fig-width: 8

    set.seed(567)
    pm <- matrix(c( .1, .11, .12,  1, 
                    .11, .5, .1,   0,
                    .12, .1, .5,  0,
                     1,  0,  0,   0), nrow=4)
    pm[lower.tri(pm)] = t(pm)[lower.tri(pm)]
    gr <- play_blocks(n = 14, size_blocks = c(4, 2, 4, 4), 
                      p_between = pm)
    V(gr)$name <- LETTERS[1:length(V(gr))]
    p <- ggraph(gr, layout = 'kk') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.07) 
    p <- p + geom_node_point(color = "tan2", size = 16) 
    p <- p + geom_node_text(aes(label = name), size = 8, color = "white")
    p <- p + theme_graph() 
    p
```

- In the matrix below, write down the cell entries for the **adjacency matrix** corresponding to the graph shown in @fig-grex1:

```{r}
    a <- matrix(as_adjacency_matrix(gr), nrow = length(V(gr)))
    rownames(a) <- LETTERS[1:length(V(gr))]
    colnames(a) <- LETTERS[1:length(V(gr))]
    diag(a) <- "----"
    kbl(a, format = "html", align = "c")  %>% 
         column_spec(1, bold = TRUE) %>% 
         column_spec (2:nrow(a), 
                      border_left = T, border_right = T) %>%
         row_spec(1:length(V(gr)), 
                  extra_css = "border-bottom: 1px solid") %>%
         kable_styling()
```

- Using the information from the previous matrix, in the matrix below, write down the cell entries for the **structural similarity matrix** corresponding to the graph shown in @fig-grex1 using the **Euclidean Distance**:

```{r}
    eudist <- function(x) {
      r <- nrow(x)
      edist <- matrix(0, r, r)
      for (i in 1: r) {
          for (j in 1: r) {
            e <- x[i, ] - x[j, ]
            e <- e^2
            e <- round(sqrt(sum(e)), 1)
            edist[i, j] <- e
          }
        }
    return(edist)
    }
    a <- matrix(as_adjacency_matrix(gr), nrow = length(V(gr)))
    a <- eudist(a)
    rownames(a) <- LETTERS[1:length(V(gr))]
    colnames(a) <- LETTERS[1:length(V(gr))]
    kbl(a, format = "html", align = "c")  %>% 
         column_spec(1, bold = TRUE) %>% 
         column_spec (2:length(V(gr)), 
                      border_left = T, border_right = T) %>%
         row_spec(1:length(V(gr)), 
                  extra_css = "border-bottom: 1px solid") %>%
         kable_styling()
```

1. Using the Information in the previous matrix, write down the set of nodes that are **structurally equivalent** (e.g., connected to the same set of neighbors) in the graph shown in @fig-grex1.
<br>
K, L, M, N

1. Out of the nodes that are not structurally equivalent, which pairs of nodes have the highest **structural similarity** in the graph shown in @fig-grex1?
<br>
A, B

1. Using the **Jaccard Similarity** metric, compute the structural similarity between nodes *B* and *D* in the graph shown in @fig-grex1.
<br>
```{r}
    m <- matrix(as_adjacency_matrix(gr), nrow = length(V(gr)))
    a <- sum(m[2, ])
    b <- sum(m[4, ])
    i <- sum(m[2, ] * m[4, ])
    J <- round(i / (a + b - i), 2)
    kbl(t(J))
```


1. Using the **Dice Similarity** metric, compute the structural similarity between nodes *C* and *D* in the graph shown in @fig-grex1.
<br>
```{r}
    a <- sum(m[3, ])
    b <- sum(m[4, ])
    i <- sum(m[3, ] * m[4, ])
    D <- round((2 * i) / (a + b), 2)
    kbl(t(D))
```


1. Using the **Cosine Similarity** metric, compute the structural similarity between nodes *A* and *C* in the graph shown in @fig-grex1.
<br>
```{r}
    a <- sum(m[1, ])
    b <- sum(m[3, ])
    i <- sum(m[1, ] * m[3, ])
    C <- round(i / sqrt(a * b), 2)
    kbl(t(C))
```

1. Write down a **node cut set** for the graph shown in @fig-grex1.
<br>
[Grade Generously]


1. Write down an **edge cut set** for the graph shown in @fig-grex1.
<br>
[Grade Generously]

1. What is the **k-connectivity** of the graph?
<br>
3 (also accept 4)

1. What is the **edge connectivity** of the graph?
<br>
3 (also accept 4)

1. What is the **pairwise k-connectivity** between nodes *E* and *G*?
<br>
4

