---
title: "Homework III: Graph Metrics"
---

```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
    require(ggraph)
    require(tidygraph)
    require(igraph)
```

```{r}
#| label: fig-grex1
#| fig-cap: "An undirected graph."
#| fig-cap-location: margin   
#| fig-height: 12
#| fig-width: 12

    set.seed(418)
    w <- 0
    while (w == 0) {
      d <- sample(1:10, 16, replace = TRUE)
      w <- as.numeric(is.graphical.degree.sequence(d))
      }
    gr <- play_degree(d, method = "simple")
    gr <- mutate(gr, name = toupper(letters[1:length(d)]))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_link(color = "steelblue", width = 1.25) 
    p <- p + geom_node_point(aes(x = x, y = y), size = 16, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 7, color = "white")
    p <- p + theme_graph() 
    p
```

Consider the graph shown in @fig-grex1:

1. What is the **order** of the graph?
<br>
<br>

1. Write down the graph's **degree sequence**:
<br>
<br>

1. What is the graph's **degree range**?
<br>
<br>

1. What is the graph's **sum of degrees**?
<br>
<br>

1. What is the **size** of the graph?
<br>
<br>

1. What is the graph's **average degree**?
<br>
<br>

1. What is the graph's **maximum size**?
<br>
<br>

1. Compute the **density** of the graph:
<br>
<br>
<br>
<br>



```{r}
#| label: fig-grex2
#| fig-cap: "A directed graph."
#| fig-cap-location: margin   
#| fig-height: 12
#| fig-width: 12

    set.seed(924)
    gr <- play_gnm(n = 13, m = 28, directed = TRUE, loops = FALSE)
    gr <- mutate(gr, name = toupper(letters[1:13]))
    p <- ggraph(gr, layout = 'auto') 
    p <- p + geom_edge_parallel(color = "steelblue", edge_width = 1.1,
                            arrow = arrow(length = unit(4, 'mm')),
                            end_cap = circle(7, 'mm'), 
                            sep = unit(8, 'mm'))    
    p <- p + geom_node_point(aes(x = x, y = y), size = 16, color = "tan2") 
    p <- p + geom_node_text(aes(label = name), size = 7, color = "white")
    p <- p + theme_graph()
    p
```

Consider the graph shown in @fig-grex2:

1. What is the graph's **out-degree sequence**?
<br>
<br>

1. What is the graph's **in-degree sequence**?
<br>
<br>

1. What is the graph's **out-degree range**?
<br>
<br>

1. What is the graph's **in-degree range**?
<br>
<br>

1. What is the graph's **sum of degrees**?
<br>
<br>

1. What is the **size** of the graph?
<br>
<br>

1. What is the graph's **maximum size**?
<br>
<br>

1. Compute the **density** of the graph:
<br>
<br>
<br>
<br>

