---
title: "Using Simple Correspondence Analysis to Present Tabular Data"
description: "Post using Correspondence Analysis to visualize tabular data with Examples."
author: "Omar Lizardo"
date: "2023-06-22"
categories: [Sociology of taste, Correspondence Analysis]
draft: false
---


```{r Setup, include=FALSE}
     knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
     library(conflicted)
     library(FactoMineR)
     library(kableExtra)
     library(sjPlot)
     library(tidyverse)
     library(wesanderson)
```

# CA and Tabular Bias
A lot of the time **Simple Correspondence Analysis** (here after CA) is presented as a complex, almost mystical technique, and its association with a relatively obscure data-analytic school in France and later Bourdieu's *Distinction* does not help. This is a pity, since CA is actually a quantitative data reduction technique that is perfect for the presenting qualitative data. Particular data from samples typical of those that ethnographers, interviewers, small-sample content-analysts, and historical sociologists deal with frequently. 

A lot of the time, these researchers will present key pieces data in tabular form, and ask the readers to use a lot of brainpower to detect patterns by staring at numbers when it would be much easier to a do a simple CA and present the categories in the rows and columns of the table as a biplot of the first two-dimensions of a simple CA. I present three examples from published work showing this "tabular bias" and show how CA could have improved things greatly (and maybe revealed some insights). Note that picked these three articles not to drag them, but because *I liked them quite a lot*, and saw potential for them to be even better than they are. 

# Example I: Journals and Theory Diagrams

```{r}
    tab <- matrix(c(72, 13, 7, 5, 3, 5, 0, 0, 0, 3, 1, 0, 1, 109, 
             27, 11, 4, 5, 19, 3, 1, 6, 11, 10, 1, 0,
             110, 34, 14, 3, 14, 24, 1, 0, 3, 9, 10, 0, 0,
             14, 11, 9, 5, 5, 5, 4, 0, 0, 2, 0, 1, 0,
             105, 57, 36, 34, 34, 24, 1, 4, 15, 3, 7, 17, 1,
             162, 13, 5, 6, 6, 9, 3, 0, 2, 6, 2, 3, 0,
             103, 35, 23, 12, 12, 20, 0, 0, 8, 3, 9, 0, 0), nrow = 7, byrow = TRUE)
    cols <- c("Arrows", "Circles", "Boxes", "Lines", "Triangles", 
              "Zig-Zags", "Legend", "Reading-Rules", "Path-Diag.",
              "Venn-Diag.", "Con-Circles")
    rows <- c("AJS", "ASR", "EJS", "EJST", "ST", "TCS", "TS")
    tab <- tab[, -c(1:2)] #dropping frist two columns
    colnames(tab) <- cols
    rownames(tab) <- rows
```

In a fascinating piece, Brett et al. (2020) study the ways sociologists use diagrams to visualize theory ideas. They select seven primarily or heavily "theoretical" journals in sociology (p. 227), and coded the "theory figures" included in a random sample of papers from each journal. The theory figure codes aimed at specifying what sort of visual languages authors used to construct the visualizations (e.g., boxes, arrows, geometrical figures, Venn diagrams and the like). The tabular data is presented as an Appendix (A) to the paper. These data are reproduced in @tbl-theoryfigs. The table has seven rows (for the journals) and eleven columns (for the theory figure motifs)

```{r}
#| label: tbl-theoryfigs
#| tbl-cap: Table of journals by theory figure characteristics.
    kbl(tab, format = "html", align = "c", padding = 2) %>% 
        #row_spec(1, bold = TRUE) %>%  
        column_spec(1, bold = TRUE) %>%  
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"))
```

The bulk of Brett et al.'s paper is concerned with specific examples, and the point is to delve into these to diagnose various problems in the ways sociologists use theory pictures (and end with some really good advice). Brett et al. were not concerned with the *relationship* between different journals and theory figure motifs, although they could have presented such data in a clear way by using a CA visualization of the table presented in the appendix. This is shown in @fig-theoryfigs.


```{r}
#| label: fig-theoryfigs
#| fig-cap: CA Visualization of the table of journals by theory figure characteristics.
#| fig-width: 8
#| fig-height: 8
    ca.res <- CA(tab)
```

The Figure shows that the distribution of theory figure motifs across theory journals is not random, something that would be hard to tell from eyeballing the numbers in @tbl-theoryfigs. Indeed, there is a pattern across journals, separating, on the main horizontal dimension, the *American Sociological Review* (ASR) and the *European Journal of Sociology* (EJS) on the one end from *Sociological Theory* (ST) at the other end. Both ASR and AJS are awash in **lines** and **path-diagrams**, while ST traffics heavily in **circles** (concentric and otherwise), **zig-zags**, **boxes**, and **Venn diagrams**. The *American Journal of Sociology* (AJS), square in the middle, balances between these two tendencies. The second dimension, on the other hand, separates the *European Journal of Social Theory* (EJST), which somehow gravitates towards **triangles**, from *Theory & Society* which likes **legends**. The **arrows** theory figure motif is neutral across journals, being placed close to the origin of the plot. 




# References
Brett, G., Silver, D., & Beelen,, K. (2020). The right tool for the job: problems and solutions in visualizing sociological theory. *Journal for the Theory of Social Behaviour*, 50(2), 223-248.

