---
title: "Bonacich Centrality and the Method of Reflections"
author: "Omar Lizardo"
draft: true
---

In a previous post, I showed how the "method of reflections" introduced by Hidalgo & Hausmann (2009)---hereafter HH---works in the case of two-mode network data. While reading that, some of you might have thought that there is a well-established network centrality metric for two-mode networks that is supposed to work just like the method of reflections: That is, defining the centrality of entities in one mode (e.g., persons) according to centralities of the entities in the other mode (e.g., groups). I refer to the Bonacich "eigenvector" centrality---see Bonacich (1991) and Faust (1997, p.) for discussion.

```{r Southern Women Data, echo=FALSE}
#| label: tbl-southern
#| tbl-cap: Southern Women Data.
#| tbl-colwidths: [2]

   library(kableExtra)
    A <- matrix(
         c(
          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 0, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 0, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0, 
          
          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 1, 1, 0, 
          
          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
          ncol = 14, byrow = TRUE)
     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
     e <- paste("E", c(1:14), sep = "")
     rownames(A) <- w
     colnames(A) <- e
     kbl(A, align = "c", format = "html", padding = 2) %>% 
        footnote(general = "Rows ordered according to the generalized blockmodeling solution of Doreian et al. (2004, Table 4).")
```

In this post I show how the Bonacich centrality fits into the reflection idea, and why this means it is a different metric from that yielded by HH's reflection method. 

Consider an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ where $|P|$ is the cardinality of the person set and $|G|$ the cardinality of the group set in the two-mode network. Each entry in the matrix $a_{pg} = 1$ if person *p* affiliates with group $g$ and zero otherwise. 

In this case, the degree centrality of people is given by:

$$
   C^R_p(1) = \sum_g a_{pg}
$$ {#eq-pc}

And for groups:

$$
   C^R_g(1) = \sum_p a_{pg}
$$ {#eq-gc}

Once we have this information, we can define a second order "Bonacich-reflection" on both the persons and groups using the formulas:

$$
   C^R_p(2) = \sum_g a_{pg}C^R_g(1)
$$ {#eq-pr2}

$$
   C^R_g(2) = \sum_p a_{pg}C^R_p(1)
$$ {#eq-gr2}

@eq-pr2 says that people are central when the sum of the number of members of the groups they belong to is large. @eq-gr2 says groups are central when the sum of the number of memberships of the people who belong to them is large. 

We can keep on going and define a third order reflection using the formulas:

$$
   C^R_p(3) = \sum_g a_{pg}C^R_g(2)
$$ {#eq-pr3}

$$
   C^R_g(3) = \sum_p a_{pg}C^R_p(2)
$$ {#eq-gr3}

@eq-pr3 says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large. @eq-pr3 says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large. 

More generally, the reflective Bonacich centralities for persons and groups are given by:

$$
   C^R_p(q) = \sum_g a_{pg}C^R_g(q-1)
$$ {#eq-pr3}

$$
   C^R_g(q) = \sum_p a_{pg}C^R_p(q-1)
$$ {#eq-gr3}

For all q > 1. 


```{r}
   source("C:/Users/Omar Lizardo/Google Drive/R/person-group-reflections/reflections.R")
     a <- reflections(A)
     
```

