---
title: "Bonacich Centrality and the Method of Reflections"
author: "Omar Lizardo"
date: "2024-05-05"
categories: [reflections, two-mode networks, centrality, eigenvectors]
draft: false
---

```{r setup, include=FALSE}
     knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
   library(kableExtra)
   source("C:/Users/Omar Lizardo/Google Drive/WORK/IN PROGRESS/NETWORKS/correspondence-analysis-two-mode-networks/Functions/sw.dat.R")
     source("C:/Users/Omar Lizardo/Google Drive/WORK/IN PROGRESS/NETWORKS/correspondence-analysis-two-mode-networks/Functions/bump.plot.R")

```

In a [previous post](https://olizardo.github.io/mysite/posts/two-mode-reflections/), I showed how the "method of reflections" introduced by Hidalgo & Hausmann (2009)---hereafter HH---works in the case of two-mode network data. While reading that, some of you might have thought that there is a well-established network centrality metric for two-mode networks that is supposed to work just like the method of reflections: That is, defining the centrality of entities in one mode (e.g., persons) according to centralities of the entities in the other mode (e.g., groups). I refer to the Bonacich "eigenvector" centrality---see Bonacich (1991) and Faust (1997) for discussion.

```{r}
#| label: tbl-sw
#| tbl-cap: Southern Women Data.

   kbl(sw.dat()[sort(rownames(sw.dat())), ], 
       align = "c", format = "html")  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))
```

In this post I show how the Bonacich centrality fits into the reflection idea, and why this means it is a different metric from that yielded by HH's reflection method. This also the difference between Bonacich and Correspondence Analysis (CA), because the HH reflection method is equivalent to CA. 

Consider an affiliation matrix $\mathbf{A}$ of dimensions $|P| \times |G|$ where $|P|$ is the cardinality of the person set and $|G|$ the cardinality of the group set in the two-mode network. Each entry in the matrix $a_{pg} = 1$ if person *p* affiliates with group $g$ and zero otherwise. An affiliation matrix of this type, representing Davis et al.'s (1941) famous Southern Women data, is shown as @tbl-sw. 

In this case, the degree centrality of people is given by:

$$
   C^B_p(1) = \sum_g a_{pg} 
$$ {#eq-pc}

And for groups:

$$
   C^B_g(1) = \sum_p a_{pg} 
$$ {#eq-gc}

The right hand sides of @eq-pc and @eq-gc are just the first order degree centralities of people and groups (Faust, 1997). People are central when they have many memberships and groups are central when they have many members.

Once we have this information, we can define a second order "Bonacich-reflection" on both the persons and groups using the formulas:

$$
   C^B_p(2) = \sum_g a_{pg}C^R_g(1) 
$$ {#eq-pr2}

$$
   C^B_g(2) = \sum_p a_{pg}C^R_p(1) 
$$ {#eq-gr2}

@eq-pr2 says that people are central when the sum of the number of members of the groups they belong to is large. @eq-gr2 says groups are central when the sum of the number of memberships of the people who belong to them is large. 

We can keep on going and define a third order reflection using the formulas:

$$
   C^B_p(3) = \sum_g a_{pg}C^R_g(2) 
$$ {#eq-pr3}

$$
   C^B_g(3) = \sum_p a_{pg}C^R_p(2) 
$$ {#eq-gr3}

@eq-pr3 says that people are central when the sum of the sum of the number of memberships held by the people who belong to the groups they belong to is large. @eq-pr3 says that groups are central when the sum of the sum of the number of members of the groups their members belong to is large. 

Once again, we can keep going and define fourth order, fifth order, and higher reflections $C^R_p(4), C^R_p(4) \ldots C^R_p(q)$.

More generally, the reflective Bonacich centralities for persons and groups are given by:

$$
   C^B_p(q) = \sum_g a_{pg}C^R_g(q-1) 
$$ {#eq-prq}

$$
   C^B_g(q) = \sum_p a_{pg}C^R_p(q-1) 
$$ {#eq-grq}

For all q > 1. 

Actually, the preceding set of equations won't quite work, because we are computing sums of sums, of sums which means that both the person and group reflective Bonacich centralities would diverge toward infinity as $q$ increases. 

To prevent this divergence and guarantee *convergence* to a set of manageable scores after a few iterations, we need to *normalize* the sum of centralities at each step $q > 1$ before feeding them to the subsequent sum at step $q+1$. We do this by dividing the computed centrality vectors for both persons and group at each step q by a *vector norm*, which in this case is the *Euclidean vector norm*, as follows:

$$
   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
$$ {#eq-prqn}

$$
   C^B_p(q) = \frac{C^B_p(q)}{||C^B_p(q)||_2}
$$ {#eq-grqn}

With the Euclidean norm for each centrality vector defined as:

$$
   ||C^B_p(q)||_2 = \sqrt{\sum_i^{|P|}(C^B_p(q)_i)^2}
$$ {#eq-peuc}

$$
   ||C^B_g(q)||_2 = \sqrt{\sum_j^{|G|}(C^B_g(q)_j)^2}
$$ {#eq-geuc}

```{r}
   bon.ref <- function(x, iter = 20) { #x is a matrix with people as rows and groups as columns iters is number of reflections
        p <- nrow(x) #number of people
        g <- ncol(x) #number of groups
        p.c <- matrix(0, p, iter) #initialize person centrality trajectories in a p X iter matrix
        g.c <- matrix(0, g, iter) #initialize group centrality trajectories matrix in a g X iter matrix
        rownames(p.c) <- rownames(x)
        rownames(g.c) <- colnames(x)
        colnames(p.c) <- paste("", c(1:iter), sep = "")
        colnames(g.c) <- paste("", c(1:iter), sep = "")
        
        p.c[, 1] <- rowSums(x) #initialize person degree centralities
        g.c[, 1] <- colSums(x) #initialize group degree centralities 
        
        k <- 1 #initializing counter
        while (k < iter) {
             m <- k + 1
             p.c[, k] <- p.c[, k]/norm(as.matrix(p.c[, k]), type = "E") #norming person centrality vector to max
             g.c[, k] <- g.c[, k]/norm(as.matrix(g.c[, k]), type = "E") #norming group centrality vector to max
             for(i in 1:p) {
                  p.c[i, m] <- sum(x[i, ] * g.c[, k])  #assign person avg. centrality groups they belong to
             } #end person loop
             for(j in 1:g) {
                  g.c[j, m] <- sum(x[, j] * p.c[, k])  #assign group avg. person centrality of people in group
             } #end group loop
             k <- k + 1 #increase counter
        } #end while loop 
        p.c[, k] <- p.c[, k]/norm(as.matrix(p.c[, k]), type = "E") #norming person centrality vector to max
        g.c[, k] <- g.c[, k]/norm(as.matrix(g.c[, k]), type = "E") #norming group centrality vector to max
        return(list(p.c = p.c, g.c = g.c))
   }
   ref.res <- bon.ref(sw.dat(), 20)
```

Let's see how that looks like with real data. The results of the reflection equations for the Southern Women data, for people up to twenty reflections, are shown in @tbl-refp. The corresponding results for groups are shown in @tbl-refg. 

```{r}
#| label: tbl-refp
#| tbl-cap: Bonacich reflections on Southern Women Data (Persons)


   kbl(round(ref.res$p.c, 2), 
       align = "c", format = "html")  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))

```

```{r}
#| label: tbl-refg
#| tbl-cap: Bonacich reflections on Southern Women Data (Groups)

   kbl(round(ref.res$g.c, 2), 
       align = "c", format = "html")  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))

```

We can see that the results for both persons and groups freeze into values after a few iterations. The resulting final ranks of persons and groups are shown in @tbl-final-1 and @tbl-final-2.

```{r}
#| label: tbl-final
#| tbl-cap: Final rankings of persons and groups according to Bonacich reflections.
#| tbl-subcap: 
#|   - Final rank of persons
#|   - Final rank of groups
#| layout-ncol: 2

   p <- sort(round(ref.res$p.c, 2)[, 20], decreasing = TRUE)
   g <- sort(round(ref.res$g.c, 2)[, 20], decreasing = TRUE)
   
   kbl(p, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))
   
   kbl(g, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))

```

The final rankings of both persons and groups should look familiar, for they are equivalent to the "dual" Bonacich centralities that can be calculated directly using methods from linear algebra as detailed in Bonacich (1991). That is, consider the standard one-mode person and group projections, which can be obtained from the original affiliation matrix as discussed in Breiger (1974). 

These are given by:

$$
B_p = AA^T
$$ {#eq-brei1}

$$
B_g = A^TA
$$ {#eq-brei1}

Where $B_p$ is the $|P| \times |P|$ person projection matrix (with off-diagonal cells indicating the number of groups two people have in common and diagonal cells indicting each person's number of memberships),  $B_g$ is the $|G| \times |G|$ group projection matrix (with off-diagonal cells indicating the number of people two groups have in common and diagonal cells indicting each group's number of members), and $A$ is the original affiliation matrix. For both persons and groups the projection matrices are shown in @tbl-proj-1 and @tbl-proj-2. 

```{r}
#| label: tbl-proj
#| tbl-cap: One mode projections for Southern Women Data.
#| tbl-subcap: 
#|   - One mode projection of persons
#|   - One mode projection of groups
#| layout-ncol: 1

   Bp <- sw.dat() %*% t(sw.dat())
   Bg <- t(sw.dat()) %*% sw.dat()
   
   kbl(Bp, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))
   
   kbl(Bg, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))

```

Bonacich shows that vectors that will order people and groups in the same way as that shown in @tbl-final, can be obtained by solving the following system of linear equations:

$$
B_p c_p = \lambda c_p
$$ {#eq-bon1}

$$
B_g c_g = \lambda c_g
$$ {#eq-bon2}

Which means that $c_p$ is an eigenvector of the projection matrix $B_p$ and $c_g$ is an eigenvector of the projection matrix $B_g$. In fact, the person and group vectors of centrality scores that will order people and groups exactly as in @tbl-final, are the leading (first) eigenvectors of each respective projection matrix (Bonacich, 1991; Faust, 1997). These are shown in @tbl-eigen. As we can see, the (absolute) value for the eigenvectors of persons and groups in @tbl-eigen is identical to that shown in @tbl-final. The reason for this is that the reflective iterative method, is actually a way of computing the eigenvectors of a matrix! (It's called the "power" method and was first discovered by von Mises)


```{r}
#| label: tbl-eigen
#| tbl-cap: Leading Eigenvectors for projection matrices of Southern Women Data.
#| tbl-subcap: 
#|   - Eigenvector scores of persons
#|   - Eigenvector scores of groups
#| layout-ncol: 2

   cp <- round(eigen(Bp)$vectors[, 1], 2)
   cg <- round(eigen(Bg)$vectors[, 1], 2)
   names(cp) <- rownames(sw.dat())
   names(cg) <- colnames(sw.dat())
   cp <- sort(cp)
   cg <- sort(cg)
   
   kbl(cp, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))
   
   kbl(cg, align = "c", format = "html", col.names = NULL)  %>% 
         column_spec(1, bold = TRUE) %>% 
         kable_styling(bootstrap_options = 
                          c("striped", "hover", "condensed", "responsive"))

```


## References

Bonacich, P. (1991). Simultaneous group and individual centralities. *Social Networks*, 13(2), 155-168.

Breiger, R. L. (1974). The duality of persons and groups. *Social Forces*, 53(2), 181-190.

Davis, A., Gardner, B. B., & Gardner, M. R. (1941). *Deep South: A social anthropological study of caste and class*. University of Chicago Press

Faust, K. (1997). Centrality in affiliation networks. *Social Networks*, 19(2), 157-191.


